---
title: "*S. pneumoniae* strain invasiveness in adults"
output: rmarkdown::html_vignette
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width = 8,
  comment = "#>"
)
```

```{r setup, echo=FALSE, warning=F, message=FALSE}
# Load dependencies
require(tidyverse)
require(magrittr)
require(rstan)
require(bridgesampling)
require(loo)
require(cowplot)
require(ggrepel)
require(xlsx)
require(gtools)
require(roxygen2)
# Load stan package
library(progressionEstimation)
#pkgbuild::compile_dll()
#roxygen2::roxygenize(package.dir = "..")

#############
# Functions #
#############

plot_progression_rate_by_type <- function(model_output_df,
                                          type = "type",
                                          unit_time = "unit time",
                                          type_name = "type",
                                          colour_col = NULL,
                                          colour_palette = NULL,
                                          use_sample_size = FALSE) {
  if (!(any(grepl("nu",colnames(model_output_df))))) {
    stop("Need to include model output in data frame for plotting")
  }
  progression_rate_values <- c("nu", "nu_lower", "nu_upper")
  if (type == "strain") {
    progression_rate_values <- paste0("secondary_", progression_rate_values)
  }
  y_label_text = paste0("Progression rate contribution (disease per carrier per ",unit_time,")")

  model_output_df %<>%
      dplyr::group_by(!!! dplyr::syms(type)) %>%
      dplyr::mutate(num_observations = sum(carriage+disease)) %>%
      dplyr::ungroup()

  if (type == "type") {
    model_output_df %<>%
      dplyr::group_by(!!! dplyr::syms(type),strain) %>%
      dplyr::slice_head(n=1) %>%
      dplyr::ungroup()
  } else if (type == "strain") {
    model_output_df %<>%
      dplyr::group_by(!!! dplyr::syms(type),type) %>%
      dplyr::slice_head(n=1) %>%
      dplyr::ungroup()
  }

  if (is.null(colour_col)) {
    if (use_sample_size) {
      base_graph <-
        ggplot(model_output_df,
               aes(x = get(!!type),
                   y = get(progression_rate_values[1]),
                   ymin = get(progression_rate_values[2]),
                   ymax = get(progression_rate_values[3]),
                   shape = num_observations))
    } else {
      base_graph <-
        ggplot(model_output_df,
               aes(x = get(!!type),
                   y = get(progression_rate_values[1]),
                   ymin = get(progression_rate_values[2]),
                   ymax = get(progression_rate_values[3])))
    }
  } else {
    if (use_sample_size) {
      base_graph <-
        ggplot(model_output_df,
               aes(x = get(!!type),
                   y = get(progression_rate_values[1]),
                   ymin = get(progression_rate_values[2]),
                   ymax = get(progression_rate_values[3]),
                   colour = get(colour_col),
                   fill = get(colour_col),
                   shape = num_observations))
    } else {
      base_graph <-
        ggplot(model_output_df,
               aes(x = get(!!type),
                   y = get(progression_rate_values[1]),
                   ymin = get(progression_rate_values[2]),
                   ymax = get(progression_rate_values[3]),
                   colour = get(colour_col),
                   fill = get(colour_col)))
    }
  }

  point_graph <-
    base_graph +
    geom_point() +
    geom_errorbar() +
    ylab(y_label_text) +
    xlab(type_name) +
    scale_y_continuous(trans ="log10") +
    theme_bw()

  if (!is.null(colour_col) & !is.null(colour_palette)) {
    point_graph <- point_graph +
      scale_colour_manual(values = colour_palette,
                          name = colour_col) +
      scale_fill_manual(values = colour_palette,
                        name = colour_col) +
      theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
            legend.position = "bottom")
  } else if (!is.null(colour_col)) {
    point_graph <- point_graph +
      theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
            legend.position = "bottom")
  } else {
    point_graph <- point_graph +
      theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
  }

  if (use_sample_size) {
    point_graph <- point_graph +
      scale_shape_binned(name = "Number of\nisolates",
                         breaks = c(5,10,25,50,100))
  }

  return(point_graph)
}

# Define serotype categories and palettes
vaccine_types <- list(
  "PCV7" = c("4","6B","9V","14","18C","19F","23F"),
  "PCV10" = c("1","5","7F"),
  "PCV13" = c("6A","3","19A"),
  "PCV15" = c("22F","33F"),
  "PCV20" = c("8","10A","11A","12F","15B/C"),
  "PPV23" = c("2","9N","17F","20")
)

vaccine_colours <- c("No PCV" = "blue",
                     "PCV7" = "red",
                     "PCV10" = "orange",
                     "PCV13" = "coral2",
                     "PCV15" = "pink",
                     "PCV20" = "purple",
                     "PPV23" = "black")

epi_colors <- c(
  "carriage" = "skyblue",
  "disease" = "cornflowerblue"
)

```

## Description and filtering of the dataset

The dataset used in this analysis is contained in the `S_pneumoniae_mixed_strain` object, which includes data from `r S_pneumoniae_mixed_strain %>% dplyr::select(study) %>% dplyr::n_distinct()` studies on `r S_pneumoniae_mixed_strain %>% dplyr::select(type) %>% dplyr::n_distinct()` serotypes. The distribution of samples between each study can be plotted, if they are ranked by the mean number of carriage and disease isolates: The total number of observations for each serotype can be plotted:

``` {r Plot distribution by serotype}


n_chains <- 2
n_iter <- 2.5e4
n_core <- 2
adapt_delta = 0.99
stepsize = 0.01
max_treedepth = 20

serotype_by_count <-
  S_pneumoniae_mixed_strain %>%
    group_by(type) %>%
    dplyr::mutate(total_count = sum(carriage)+sum(disease)) %>%
    dplyr::select(type,total_count) %>%
    dplyr::distinct() %>%
    dplyr::arrange(desc(total_count)) %>%
    dplyr::select(type) %>%
    dplyr::pull()

S_pneumoniae_mixed_strain %<>%
  dplyr::mutate(type = factor(type, levels = serotype_by_count)) %>%
  dplyr::mutate(study = factor(study))

mixed_strain_serotype_distribution_plot <-
  ggplot(S_pneumoniae_mixed_strain %>%
           tidyr::pivot_longer(c(carriage,disease),
                               names_to = "Sample type",
                               values_to = "count"),
         aes(x = type,
             y = count,
             colour = `Sample type`,
             fill = `Sample type`)) +
    geom_col() +
    xlab("Serotype") +
    theme_bw() + 
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
    scale_fill_manual(values = epi_colors) +
    scale_colour_manual(values = epi_colors)

ggsave(mixed_strain_serotype_distribution_plot,
       file = "mixed_strain_serotype_distribution_plot.png",
       height = 6,
       width = 6
)

```

The study-adjusted type-specific model can be applied to the overall dataset:

``` {r Plot invasiveness estimates from best-fitting model}

full_mixed_serotype_and_strain_invasiveness <- 
  progressionEstimation::process_input_data(S_pneumoniae_mixed_strain,
                                            use_strain = TRUE,
                                            condense = FALSE)

full_adjusted_serotype_specific_strain_modified_poisson_fit <- 
  progressionEstimation::fit_progression_rate_model(full_mixed_serotype_and_strain_invasiveness,
                                                    type_specific = TRUE,
                                                    location_adjustment = FALSE,
                                                    strain_as_secondary_type = TRUE,
                                                    stat_model = "poisson",
                                                    model_description = "serotype_specific_strain_modified_poisson",
                                                    num_chains = n_chains,
                                                    num_iter = n_iter,
                                                    num_cores = n_core,
                                                    adapt_delta_value = adapt_delta,
                                                    stepsize_value = stepsize,
                                                    max_treedepth_value = max_treedepth)

rhat_plot <- plot(full_adjusted_serotype_specific_strain_modified_poisson_fit, plotfun = "rhat", binwidth = 0.00005)

trace_plot <- rstan::traceplot(full_adjusted_serotype_specific_strain_modified_poisson_fit, pars = "lp__") + theme(axis.text.x = element_text(angle = 90))

(full_strain_and_serotype_validation_plot <-
  cowplot::plot_grid(plotlist = list(rhat_plot,trace_plot),
                     nrow = 1,
                     ncol = 2,
                     labels = "AUTO")
)

ggsave(full_strain_and_serotype_validation_plot,
       file = "full_mixed_strain_and_serotype_validation_plot.png",
       width = 7,
       height = 5
)

```

``` {r Plot invasiveness across all serotypes}

# Annotate serotypes by status
best_fitting_output <- 
  progressionEstimation::process_progression_rate_model_output(full_adjusted_serotype_specific_strain_modified_poisson_fit,
                                                               S_pneumoniae_mixed_strain,
                                                               strain_as_secondary_type = TRUE,
                                                               condense = FALSE) %>%
    dplyr::mutate("type" = S_pneumoniae_mixed_strain$type) %>%
    dplyr::mutate("classification" = dplyr::case_when(
        type %in% vaccine_types[["PCV7"]] ~ "PCV7",
        type %in% vaccine_types[["PCV10"]] ~ "PCV10",
        type %in% vaccine_types[["PCV13"]] ~ "PCV13",
        type %in% vaccine_types[["PCV15"]] ~ "PCV15",
        type %in% vaccine_types[["PCV20"]] ~ "PCV20",
        type %in% vaccine_types[["PPV23"]] ~ "PPV23",
        TRUE ~ "No PCV"
      )
    ) %>%
    dplyr::mutate(classification = factor(classification, levels = c("PCV7",
                                                                     "PCV10",
                                                                     "PCV13",
                                                                     "PCV15",
                                                                     "PCV20",
                                                                     "PPV23",
                                                                     "No PCV"))) %>%
    tidyr::unite(combined,type,strain,remove = FALSE) %>%
    dplyr::group_by(combined) %>%
    dplyr::mutate(num_observations = sum(carriage+disease)) %>%
    dplyr::ungroup() %>%
    dplyr::filter(num_observations >= 10) %>%
    dplyr::group_by(strain) %>%
    dplyr::mutate(num_serotypes = dplyr::n_distinct(type)) %>%
    dplyr::ungroup() %>%
    dplyr::group_by(type) %>%
    dplyr::mutate(num_strains = dplyr::n_distinct(strain)) %>%
    dplyr::ungroup()

# Function needed if processing combined serotype/strain assignments
mixedrank = function(x) order(gtools::mixedorder(as.character(x)))

# Order serotypes
serotype_levels <-
  best_fitting_output %>%
    dplyr::select(type,classification) %>%
    dplyr::distinct() %>%
    dplyr::group_by(classification) %>%
    dplyr::arrange(mixedrank(type),
                   .by_group = TRUE) %>%
    dplyr::ungroup() %>%
    dplyr::select(type) %>%
    dplyr::mutate(type = as.character(type)) %>%
    dplyr::pull()

strain_levels <- gtools::mixedsort(best_fitting_output$strain)

serotypes_by_invasiveness <-
  best_fitting_output %>%
    dplyr::select(type,nu) %>%
    dplyr::distinct() %>%
    dplyr::arrange(desc(nu)) %>%
    dplyr::select(type) %>%
    dplyr::mutate(type = as.character(type)) %>%
    dplyr::pull()

strains_by_invasiveness <-
  best_fitting_output %>%
    dplyr::select(strain,secondary_nu) %>%
    dplyr::distinct() %>%
    dplyr::arrange(desc(secondary_nu)) %>%
    dplyr::select(strain) %>%
    dplyr::mutate(strain = as.character(strain)) %>%
    dplyr::pull()

best_fitting_output %<>%
  dplyr::mutate(combined = factor(strain, levels = strains_by_invasiveness))

# Plot
(full_invasiveness_plot <-
  plot_progression_rate_by_type(best_fitting_output %>%
                                                  dplyr::filter(num_strains > 1),
                                                type = "strain",
                                                unit_time = "year",
                                                type_name = "Strain",
                                                colour_col = "classification",
                                                colour_palette = vaccine_colours,
                                                use_sample_size = TRUE) +
                                                facet_wrap(.~strain,
                                                           scales = 'free_x'))

ggsave(full_invasiveness_plot,
       file = "full_mixed_strain_and_serotype_invasiveness_plot.png",
       width = 20,
       height = 10
)

```



``` {r Plot only larger sample size}

# Annotate sample size
filtered_best_fitting_output <-
  best_fitting_output %>%
    dplyr::group_by(combined) %>%
    dplyr::mutate(num_observations = sum(carriage+disease)) %>%
    dplyr::ungroup() %>%
    dplyr::filter(num_observations >= 10) %>%
    dplyr::group_by(strain) %>%
    dplyr::mutate(num_serotypes = dplyr::n_distinct(type)) %>%
    dplyr::ungroup() %>%
    dplyr::group_by(type) %>%
    dplyr::mutate(num_strains = dplyr::n_distinct(strain)) %>%
    dplyr::ungroup()

# Plot
(filtered_invasiveness_plot <-
  plot_progression_rate_by_type(filtered_best_fitting_output %>%
                                  dplyr::mutate(strain = factor(strain, levels = strains_by_invasiveness)) %>%
                                  dplyr::mutate(type = factor(type, levels = serotype_levels)) %>%
                                                  dplyr::filter(num_strains > 1),
                                                type = "strain",
                                                unit_time = "year",
                                                type_name = "Strain",
                                                colour_col = "classification",
                                                colour_palette = vaccine_colours,
                                                use_sample_size = TRUE) +
                                                facet_wrap(.~type,
                                                           scales = 'free_x'))

ggsave(filtered_invasiveness_plot,
       file = "filtered_mixed_strain_and_serotype_invasiveness_plot.png",
       width = 20,
       height = 10
)

```
``` {r Joint plot for mixed strain invasiveness}

# Plot
(filtered_sero_invasiveness_plot <-
  plot_progression_rate_by_type(filtered_best_fitting_output %>%
                                    dplyr::filter(num_serotypes > 1) %>%
                                    dplyr::mutate(strain = factor(strain, levels = strains_by_invasiveness)) %>%
                                    dplyr::mutate(type = factor(type, levels = serotypes_by_invasiveness)),
                                                type = "type",
                                                unit_time = "year",
                                                type_name = "Serotype",
                                                colour_col = "classification",
                                                colour_palette = vaccine_colours,
                                                use_sample_size = TRUE) +
                                                facet_wrap(.~strain,
                                                           scales = 'free_x'))

(combined_mixed_strain_invasiveness_plot <-
  cowplot::plot_grid(plotlist = list(filtered_sero_invasiveness_plot +
                                       ylab("Progression rate contribution\n(disease per carrier per year)") +
                                       theme(legend.position = "none"),
                                     filtered_invasiveness_plot +
                                       ylab("Progression rate contribution\n(disease per carrier per year)")),
                     labels = "AUTO",
                     nrow = 2,
                     rel_heights = c(0.3,0.7)))

ggsave(combined_mixed_strain_invasiveness_plot,
       file = "combined_mixed_strain_invasiveness_plot.png",
       height = 12,
       width = 9)
```
