---
title: "s-pneumoniae-serotype-invasiveness-adults"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{s-pneumoniae-serotype-invasiveness}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width = 8,
  comment = "#>"
)
options(rmarkdown.html_vignette.check_title = FALSE)
```

```{r setup}
# Load dependencies
require(tidyverse)
require(magrittr)
require(rstan)
require(bridgesampling)
require(loo)
require(cowplot)
require(ggrepel)
require(xlsx)

# Load stan package
library(progressionEstimation)
pkgbuild::compile_dll()
roxygen2::roxygenize(package.dir = "..")

# Define serotype categories and palettes
vaccine_types <- list(
  "PCV7" = c("4","6B","9V","14","18C","19F","23F"),
  "PCV10" = c("1","5","7F"),
  "PCV13" = c("6A","3","19A"),
  "PCV15" = c("22F","33F"),
  "PCV20" = c("8","10A","11A","12F","15B/C"),
  "PPV23" = c("2","9N","17F","20")
)

vaccine_colours <- c("No PCV" = "blue",
                     "PCV7" = "red",
                     "PCV10" = "orange",
                     "PCV13" = "coral2",
                     "PCV15" = "pink",
                     "PCV20" = "purple",
                     "PPV23" = "black")

epi_colors <- c(
  "carriage" = "skyblue",
  "disease" = "cornflowerblue"
)

```

## Description and filtering of the dataset

The dataset used in this analysis is contained in the `S_pneumoniae_adult_serotype` object, which includes data from `r S_pneumoniae_adult_serotype %>% dplyr::select(study) %>% dplyr::n_distinct()` studies on `r S_pneumoniae_adult_serotype %>% dplyr::select(type) %>% dplyr::n_distinct()` serotypes. The distribution of samples between each study can be plotted, if they are ranked by the mean number of carriage and disease isolates:

``` {r Plot distribution by study, include = FALSE}

studies_by_size <-
  S_pneumoniae_adult_serotype %>%
    group_by(study) %>%
    dplyr::mutate(total_count = sum(carriage)+sum(disease)) %>%
    dplyr::select(study,total_count) %>%
    dplyr::distinct() %>%
    dplyr::arrange(desc(total_count)) %>%
    dplyr::select(study) %>%
    dplyr::pull()

S_pneumoniae_adult_serotype %<>%
  dplyr::mutate(study = factor(study, levels = studies_by_size))

full_epi_data_plot <-
  ggplot(S_pneumoniae_adult_serotype %>%
           tidyr::pivot_longer(c(carriage,disease),
                               names_to = "Sample type",
                               values_to = "count"),
         aes(x = study,
             y = count,
             colour = `Sample type`,
             fill = `Sample type`)) +
    geom_col() +
    theme_bw() + 
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
    scale_fill_manual(values = epi_colors) +
    scale_colour_manual(values = epi_colors)

```

The total number of observations for each serotype can be similarly plotted:

``` {r Plot distribution by serotype, include = FALSE}

serotype_by_count <-
  S_pneumoniae_adult_serotype %>%
    group_by(type) %>%
    dplyr::mutate(total_count = sum(carriage)+sum(disease)) %>%
    dplyr::select(type,total_count) %>%
    dplyr::distinct() %>%
    dplyr::arrange(desc(total_count)) %>%
    dplyr::select(type) %>%
    dplyr::pull()

S_pneumoniae_adult_serotype %<>%
  dplyr::mutate(type = factor(type, levels = serotype_by_count))

serotype_frequency_plot <-
  ggplot(S_pneumoniae_adult_serotype %>%
           tidyr::pivot_longer(c(carriage,disease),
                               names_to = "Sample type",
                               values_to = "count"),
         aes(x = type,
             y = count,
             colour = `Sample type`,
             fill = `Sample type`)) +
    geom_col() +
    xlab("Serotype") +
    theme_bw() + 
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
    scale_fill_manual(values = epi_colors) +
    scale_colour_manual(values = epi_colors)

(combined_full_epi_data_plot <-
  cowplot::plot_grid(plotlist = list(full_epi_data_plot, serotype_frequency_plot),
                     nrow = 2,
                     ncol = 1,
                     labels = "AUTO")
)

ggsave(combined_full_epi_data_plot,
       file = "full_epi_data_plot.png",
       height = 12,
       width = 8
)

```

As many of these serotypes are rare, to facilitate model fitting, we exclude any serotype for which the maximum number of observations is below five:

``` {r Plot distribution by serotype in filtered dataset, include = FALSE}

threshold_count <- 5

filtered_S_pneumoniae_adult_serotype <-
  S_pneumoniae_adult_serotype %>%
    dplyr::filter(carriage >= threshold_count | disease >= threshold_count)
# %>%
#     dplyr::group_by(type)%>%
#     dplyr::mutate(study_count = n()) %>%
#     dplyr::ungroup() %>%
#     dplyr::filter(study_count > threshold_count) %>%
#     dplyr::group_by(study) %>%
#     dplyr::mutate(serotype_count = n()) %>%
#     dplyr::ungroup() %>%
#     dplyr::filter(serotype_count > threshold_count) %>%
#     dplyr::mutate(type = factor(type))

filtered_S_pneumoniae_adult_serotype %<>%
    dplyr::mutate(study = factor(study, levels = studies_by_size[studies_by_size %in% unique(filtered_S_pneumoniae_adult_serotype$study)]))

filtered_S_pneumoniae_adult_serotype %<>%
    dplyr::mutate(type = factor(type,
                                          levels = serotype_by_count[serotype_by_count %in% unique(filtered_S_pneumoniae_adult_serotype$type)]))

adult_serotype_invasiveness <- 
  progressionEstimation::process_input_data(filtered_S_pneumoniae_adult_serotype)

filtered_epi_data_plot <-
    ggplot(filtered_S_pneumoniae_adult_serotype %>%
           tidyr::pivot_longer(c(carriage,disease),
                               names_to = "Sample type",
                               values_to = "count"),
         aes(x = study,
             y = count,
             colour = `Sample type`,
             fill = `Sample type`)) +
    geom_col() +
    theme_bw() + 
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
    scale_fill_manual(values = epi_colors) +
    scale_colour_manual(values = epi_colors)

filtered_serotype_frequency_plot <-
  ggplot(filtered_S_pneumoniae_adult_serotype %>%
           tidyr::pivot_longer(c(carriage,disease),
                               names_to = "Sample type",
                               values_to = "count"),
         aes(x = type,
             y = count,
             colour = `Sample type`,
             fill = `Sample type`)) +
    geom_col() +
    xlab("Serotype") +
    theme_bw() + 
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
    scale_fill_manual(values = epi_colors) +
    scale_colour_manual(values = epi_colors)

(combined_filtered_epi_data_plot <-
  cowplot::plot_grid(plotlist = list(filtered_epi_data_plot, filtered_serotype_frequency_plot),
                     nrow = 2,
                     ncol = 1,
                     labels = "AUTO")
)

ggsave(combined_filtered_epi_data_plot,
       file = "filtered_epi_data_plot.png",
       height = 12,
       width = 8
)

```

This filtered dataset comprises `r S_pneumoniae_adult_serotype %>% dplyr::select(type) %>% dplyr::n_distinct()` serotypes.

## Fitting the statistical models

The null model, in which each serotyoe $j$ causes invasive disease at the same rate in each location $i$, can be fitted using a Poisson distribution of disease isolates (**null Poisson** model):

$$d_{i,j} \sim Pois(\nu\rho_{i,j}N_{i}t_{i})$$

``` {r Null Poisson model fit, include = FALSE}

n_chains <- 2
n_iter <- 2.5e4
n_core <- 2

null_poisson_fit <- 
  progressionEstimation::fit_progression_rate_model(adult_serotype_invasiveness,
                                                    type_specific = FALSE,
                                                    location_adjustment = FALSE,
                                                    stat_model = "poisson",
                                                    num_chains = n_chains,
                                                    num_iter = n_iter,
                                                    num_cores = n_core)

```

The null model can be fitted using a negative binomial distribution of disease isolates (**null negative binomial** model):

$$d_{i,j} \sim NegBin(\nu\rho_{i,j}N_{i}t_{i},\phi)$$

``` {r Null negative binomial model fit, include = FALSE}

null_negbin_fit <- 
  progressionEstimation::fit_progression_rate_model(adult_serotype_invasiveness,
                                                    type_specific = FALSE,
                                                    location_adjustment = FALSE,
                                                    stat_model = "negbin",
                                                    num_chains = n_chains,
                                                    num_iter = n_iter,
                                                    num_cores = n_core)

```

The Poisson model can be modified by allowing each serotype to have its own invasiveness, $\nu_{j}$ (**type-specific Poisson** model):

$$d_{i,j} \sim Pois(\nu_{j}\rho_{i,j}N_{i}t_{i})$$
``` {r Type-specific Poisson model fit, include = FALSE}

type_specific_poisson_fit <- 
  progressionEstimation::fit_progression_rate_model(adult_serotype_invasiveness,
                                                    type_specific = TRUE,
                                                    location_adjustment = FALSE,
                                                    stat_model = "poisson",
                                                    num_chains = n_chains,
                                                    num_iter = n_iter,
                                                    num_cores = n_core)

```

The negative binomial model can be modified in the same way (**type-specific negative binomial** model):

$$d_{i,j} \sim NegBin(\nu_{j}\rho_{i,j}N_{i}t_{i},\phi)$$

``` {r Type-specific negative binomial model fit, include = FALSE}

type_specific_negbin_fit <- 
  progressionEstimation::fit_progression_rate_model(adult_serotype_invasiveness,
                                                    type_specific = TRUE,
                                                    location_adjustment = FALSE,
                                                    stat_model = "negbin",
                                                    num_chains = n_chains,
                                                    num_iter = n_iter,
                                                    num_cores = n_core)

```

Alternatively, the Poisson model can be modified by allowing invasiveness to be adjusted between locations by a scale factor, $\delta_{i}$. In this model, all serotypes still have the same invasiveness (**adjusted Poisson** model):

$$d_{i,j} \sim Poisson(\delta_{i}\nu\rho_{i,j}N_{i}t_{i})$$

``` {r Adjusted Poisson model fit, include = FALSE}

adjusted_poisson_fit <- 
  progressionEstimation::fit_progression_rate_model(adult_serotype_invasiveness,
                                                    type_specific = FALSE,
                                                    location_adjustment = TRUE,
                                                    stat_model = "poisson",
                                                    num_chains = n_chains,
                                                    num_iter = n_iter,
                                                    num_cores = n_core)

```

The negative binomial model can again be modified in the same way (**adjusted negative binomial** model):

$$d_{i,j} \sim NegBin(\delta_{i}\nu\rho_{i,j}N_{i}t_{i},\phi)$$

``` {r Adjusted negative binomial model fit, include = FALSE}

adjusted_negbin_fit <- 
  progressionEstimation::fit_progression_rate_model(adult_serotype_invasiveness,
                                                    type_specific = FALSE,
                                                    location_adjustment = TRUE,
                                                    stat_model = "negbin",
                                                    num_chains = n_chains,
                                                    num_iter = n_iter,
                                                    num_cores = n_core)

```


Both modifications can be combined in adjusted type-specific model fits. This can be used with the Poisson model of disease isolate counts (**adjusted type-specific Poisson** model):

$$d_{i,j} \sim Poisson(\delta_{i}\nu_{j}\rho_{i,j}N_{i}t_{i})$$

``` {r Adjusted type-specific Poisson model fit, include = FALSE}

adjusted_type_specific_poisson_fit <- 
  progressionEstimation::fit_progression_rate_model(adult_serotype_invasiveness,
                                                    type_specific = TRUE,
                                                    location_adjustment = TRUE,
                                                    stat_model = "poisson",
                                                    num_chains = n_chains,
                                                    num_iter = n_iter,
                                                    num_cores = n_core)

```

This can also be used with the negative binomial model (**adjusted type-specific negative binomial** model):

$$d_{i,j} \sim NegBin(\delta_{i}\nu_{j}\rho_{i,j}N_{i}t_{i},\phi)$$

``` {r Adjusted type-specific negative binomial model fit, include = FALSE}

adjusted_type_specific_negbin_fit <- 
  progressionEstimation::fit_progression_rate_model(adult_serotype_invasiveness,
                                                    type_specific = TRUE,
                                                    location_adjustment = TRUE,
                                                    stat_model = "negbin",
                                                    num_chains = n_chains,
                                                    num_iter = n_iter,
                                                    num_cores = n_core)

```

To check the models have fitted correctly, we can plot the $\hat{R}$ values across MCMCs:

``` {r Plot r hat values}

fit_list <- list(null_poisson_fit,
                 null_negbin_fit,
                 type_specific_poisson_fit,
                 type_specific_negbin_fit,
                 adjusted_poisson_fit,
                 adjusted_negbin_fit,
                 adjusted_type_specific_poisson_fit,
                 adjusted_type_specific_negbin_fit)

rhat_plots <- lapply(fit_list, plot, plotfun = "rhat", binwidth = 0.00005)

cowplot::plot_grid(plotlist = rhat_plots, ncol = 4, nrow = 2, labels = "AUTO")

```

Convergence can also be checked using trace plots of the log likelihood values:

``` {r Plot log likelihood traces}

plot_lp <- function(fit) {
  pl <- rstan::traceplot(fit, pars = "lp__")
  no_leg_pl <- pl + theme(legend.position = "none")
  return(no_leg_pl)
}

trace_plots <- lapply(fit_list, plot_lp)

cowplot::plot_grid(plotlist = trace_plots, ncol = 4, nrow = 2, labels = "AUTO")

```

## Compare model observations and predictions

For each model, the output can be processed, and the observations plotted against each model's predictions.

``` {r Process model outputs and plot predictions, fig.height = 18}

model_outputs <- lapply(fit_list,
                        progressionEstimation::process_progression_rate_model_output,
                        filtered_S_pneumoniae_adult_serotype)

prediction_plots <- lapply(model_outputs,
                           progressionEstimation::plot_case_carrier_predictions,
                           legend = FALSE)

no_leg_plot <- cowplot::plot_grid(plotlist = prediction_plots, ncol = 1, nrow = 8, labels = "AUTO")

prediction_plot_legend <- progressionEstimation::plot_case_carrier_predictions(model_outputs[[1]], just_legend = TRUE)

cowplot::plot_grid(no_leg_plot, prediction_plot_legend, ncol = 1, rel_heights = c(0.9,0.1))

```

The deviation of the predicted and observed values can be simply evaluated using the root mean square error:

``` {r Calculate RMSE}

obs_pred_df <-
  dplyr::bind_rows(model_outputs) %>%
  dplyr::select(model_name,carriage_rmse,disease_rmse,phi,phi_lower,phi_upper) %>%
  tidyr::pivot_longer(c(carriage_rmse,disease_rmse), names_to = "Observation", values_to = "RMSE") %>%
  dplyr::mutate(model_name = factor(model_name,
                                    levels = c(
                                      "null_poisson",
                                      "null_negbin",
                                      "type_specific_poisson",
                                      "type_specific_negbin",
                                      "adjusted_null_poisson",
                                      "adjusted_null_negbin",
                                      "adjusted_type_specific_poisson",
                                      "adjusted_type_specific_negbin"
                                      )
                                    )
                ) %>%
  dplyr::mutate(Observation = dplyr::case_when(
                    Observation == "carriage_rmse" ~ "Carriage",
                    Observation == "disease_rmse" ~ "Disease"
                  )
                )

ggplot(obs_pred_df, aes(x = model_name, y = RMSE)) +
  #geom_boxplot(outlier.shape = NA) +
  geom_point(position = position_jitter(width = 0.25),
             alpha = 0.25,
             cex = 0.75,
             colour = "blue",
             pch = 4) +
  geom_violin(fill = NA) +
  scale_y_continuous(trans="log10") +
  theme_bw() +
  facet_grid(Observation ~ .) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

```

The fits of the negative binomial can be evaluated through plotting the estimates of the precision parameter, $\phi$. As $\phi$ increases, the variance of the negative binomial distribution decreases, and the negative binomal model becomes more similar to the Poisson model.

``` {r Plot precision parameters for negative binomial models, fig.width = 6}

ggplot(obs_pred_df %>%
        dplyr::filter(!is.na(phi)) %>%
        dplyr::group_by(model_name) %>%
        dplyr::slice(1) %>%
        dplyr::ungroup(),
      aes(x = model_name,
          y = phi,
          ymin = phi_lower,
          ymax = phi_upper)) +
  geom_point() +
  geom_errorbar(width = 0.25) +
  theme_bw() +
  ylab(expression(phi)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

```

## Run model comparison with leave-one-out cross-validation

The model likelihoods can be compared using leave-one-out cross-validation:

``` {r Comparison with LOO-CV using full log likelihood, include = FALSE}

loo_comparison <- progressionEstimation::compare_model_fits_with_loo(fit_list,
                                                                     use_moments = FALSE,
                                                                     num_threads = n_core)

loo_comparison %>%
  kableExtra::kable()  %>%
  kableExtra::kable_styling(latex_options = "scale_down")

```

The Pareto distribution k values are too high for robust estimates of the ELPD. This reflects many of the points being 

``` {r Comparison with LOO-CV using disease distribution log likelihood, include = FALSE}

disease_loo_comparison <- progressionEstimation::compare_model_fits_with_loo(fit_list,
                                                                     log_lik_param = "disease_log_lik",
                                                                     use_moments = FALSE,
                                                                     num_threads = n_core)

disease_loo_comparison %>%
  kableExtra::kable()  %>%
  kableExtra::kable_styling(latex_options = "scale_down")

```

## Run model comparison with Bayes factors

As the models can instead be compared with Bayes factors:

``` {r Comparison with BF, include = FALSE}

bf_comparison <- progressionEstimation::compare_model_fits_with_bf(fit_list,
                                                                   num_iter = n_iter,
                                                                   num_threads = n_core)
bf_comparison %>%
  kableExtra::kable()  %>%
  kableExtra::kable_styling(latex_options = "scale_down")

```

## Plot serotype invasiveness

The best-fitting model can then be applied to the overall dataset.

``` {r Plot invasiveness estimates from best-fitting model}

full_adult_serotype_invasiveness <- 
  progressionEstimation::process_input_data(
    S_pneumoniae_adult_serotype %<>%
      dplyr::mutate(study = factor(study, levels = studies_by_size)
      )
    )

full_adjusted_type_specific_poisson_fit <- 
  progressionEstimation::fit_progression_rate_model(full_adult_serotype_invasiveness,
                                                    type_specific = TRUE,
                                                    location_adjustment = TRUE,
                                                    stat_model = "poisson",
                                                    num_chains = n_chains,
                                                    num_iter = n_iter,
                                                    num_cores = n_core)

rhat_plot <- plot(full_adjusted_type_specific_poisson_fit, plotfun = "rhat", binwidth = 0.00005)

trace_plot <- rstan::traceplot(full_adjusted_type_specific_poisson_fit, pars = "lp__")

cowplot::plot_grid(plotlist = list(rhat_plot,trace_plot),
                   nrow = 1,
                   ncol = 2,
                   labels = "AUTO")

```

``` {r Plot invasiveness across all serotypes}

mixedrank = function(x) order(gtools::mixedorder(as.character(x)))

# Annotate serotypes by status
best_fitting_output <- 
  progressionEstimation::process_progression_rate_model_output(full_adjusted_type_specific_poisson_fit,
                                                               S_pneumoniae_adult_serotype) %>%
    dplyr::mutate("classification" = dplyr::case_when(
        type %in% vaccine_types[["PCV7"]] ~ "PCV7",
        type %in% vaccine_types[["PCV10"]] ~ "PCV10",
        type %in% vaccine_types[["PCV13"]] ~ "PCV13",
        type %in% vaccine_types[["PCV15"]] ~ "PCV15",
        type %in% vaccine_types[["PCV20"]] ~ "PCV20",
        type %in% vaccine_types[["PPV23"]] ~ "PPV23",
        TRUE ~ "No PCV"
      )
    ) %>%
    dplyr::mutate(classification = factor(classification, levels = c("PCV7",
                                                                     "PCV10",
                                                                     "PCV13",
                                                                     "PCV15",
                                                                     "PCV20",
                                                                     "PPV23",
                                                                     "No PCV")))

# Order serotypes
serotype_levels <-
  best_fitting_output %>%
    dplyr::select(type,classification) %>%
    dplyr::distinct() %>%
    group_by(classification) %>%
    dplyr::arrange(mixedrank(type),
                   .by_group = TRUE) %>%
    ungroup() %>%
    dplyr::select(type) %>%
    dplyr::pull()

best_fitting_output %<>%
  dplyr::mutate(type = factor(type, levels = serotype_levels))

# Plot
(invasiveness_plot <-
  progressionEstimation::plot_progression_rates(best_fitting_output,
                                                unit_time = "year",
                                                type_name = "Serotype",
                                                colour_col = "classification",
                                                colour_palette = vaccine_colours,
                                                use_sample_size = TRUE))

ggsave(invasiveness_plot,
       file = "full_serotype_invasiveness_plot.png",
       height = 8,
       width = 8)

```

## Plot study-specific scale factors

``` {r Plot study-specific scale factor estimates from best-fitting model}

(scale_factor_plot <-
  progressionEstimation::plot_study_scale_factors(best_fitting_output))

```

## Compare a model fit that account for PCVs protecting against disease

If PCVs provide further protection against disease, relative to carriage, then their invasiveness may reduce post-PCV. Therefore we can treat them as different serotypes pre- and post-PCV.


``` {r Adjust invasiveness post-PCV}

pv_S_pneumoniae_adult_serotype <-
  S_pneumoniae_adult_serotype %>%
    dplyr::mutate(type = dplyr::case_when(
                      type %in% vaccine_types[["PCV7"]] & grepl("post.PCV", study) ~ paste0(type,"_pv"),
                      type == "6A" & grepl("post.PCV", study) ~ paste0(type,"_pv"),
                      type %in% vaccine_types[["PCV10"]] & grepl("post.PCV10", study) ~ paste0(type,"_pv"),
                      type %in% vaccine_types[["PCV10"]] & grepl("post.PCV13", study) ~ paste0(type,"_pv"),
                      type %in% vaccine_types[["PCV13"]] & grepl("post.PCV13", study) ~ paste0(type,"_pv"),
                      type == "6C" & grepl("post.PCV13", study) ~ paste0(type,"_pv"),
                      TRUE ~ as.character(type)
                    )
                  ) %>%
  dplyr::mutate(type = factor(type))

pv_S_pneumoniae_adult_serotype %<>%
  dplyr::mutate(study = factor(study, levels = studies_by_size))

pv_full_adult_serotype_invasiveness <- 
  progressionEstimation::process_input_data(pv_S_pneumoniae_adult_serotype)

pv_full_adjusted_type_specific_poisson_fit <- 
  progressionEstimation::fit_progression_rate_model(pv_full_adult_serotype_invasiveness,
                                                    type_specific = TRUE,
                                                    location_adjustment = TRUE,
                                                    stat_model = "poisson",
                                                    num_chains = n_chains,
                                                    num_iter = n_iter,
                                                    num_cores = n_core)

rhat_plot <- plot(pv_full_adjusted_type_specific_poisson_fit, plotfun = "rhat", binwidth = 0.00005)

trace_plot <- rstan::traceplot(pv_full_adjusted_type_specific_poisson_fit, pars = "lp__")

cowplot::plot_grid(plotlist = list(rhat_plot,trace_plot),
                   nrow = 1,
                   ncol = 2,
                   labels = "AUTO")
```

``` {r Analyse the pv output}

pv_best_fitting_output <- 
  progressionEstimation::process_progression_rate_model_output(
    pv_full_adjusted_type_specific_poisson_fit,                                                 pv_S_pneumoniae_adult_serotype) %>%
      dplyr::mutate("classification" = dplyr::case_when(
          gsub("_pv","",type) %in% vaccine_types[["PCV7"]] ~ "PCV7",
          gsub("_pv","",type) %in% vaccine_types[["PCV10"]] ~ "PCV10",
          gsub("_pv","",type) %in% vaccine_types[["PCV13"]] ~ "PCV13",
          gsub("_pv","",type) %in% vaccine_types[["PCV15"]] ~ "PCV15",
          gsub("_pv","",type) %in% vaccine_types[["PCV20"]] ~ "PCV20",
          gsub("_pv","",type) %in% vaccine_types[["PPV23"]] ~ "PPV23",
          TRUE ~ "No PCV"
        )
      ) %>%
      dplyr::mutate(classification = factor(classification, levels = c("PCV7",
                                                                       "PCV10",
                                                                       "PCV13",
                                                                       "PCV15",
                                                                       "PCV20",
                                                                       "PPV23",
                                                                       "No PCV")))

# Order serotypes
pv_serotype_levels <-
  pv_best_fitting_output %>%
    dplyr::select(type,classification) %>%
    dplyr::distinct() %>%
    group_by(classification) %>%
    dplyr::arrange(mixedrank(type),
                   .by_group = TRUE) %>%
    ungroup() %>%
    dplyr::select(type) %>%
    dplyr::pull()

pv_best_fitting_output %<>%
  dplyr::mutate(type = factor(type, levels = pv_serotype_levels))

# Plot
(pv_invasiveness_plot <-
  progressionEstimation::plot_progression_rates(pv_best_fitting_output,
                                                unit_time = "year",
                                                type_name = "Serotype",
                                                colour_col = "classification",
                                                colour_palette = vaccine_colours,
                                                use_sample_size = TRUE))

(joint_invasiveness_plot <-
    cowplot::plot_grid(plotlist = list(invasiveness_plot, pv_invasiveness_plot),
                       nrow = 2,
                       ncol = 1,
                       labels = "AUTO"
                       )
)

ggsave(joint_invasiveness_plot,
       file = "join_adult_serotype_plot.png",
       height = 10,
       width = 8)

```

``` {r Plot study-specific scale factor estimates from best-fitting model, adjust for vaccine}

(pv_scale_factor_plot <-
  progressionEstimation::plot_study_scale_factors(pv_best_fitting_output))

```

## Comparison with odds ratio calculation

``` {r Calculate odds ratios}

require(ggpubr)
require(metafor)

get_serotype_stats <- function(serotype, df) {
  sero_df <-
    df %>%
      tidyr::pivot_longer(c(carriage,disease),
                          names_to = "clinical_manifestation",
                          values_to = "observation") %>%
      dplyr::group_by(study) %>%
      dplyr::mutate(matching_type = dplyr::if_else(type == serotype,1,0)) %>%
      dplyr::mutate(matching_type_disease =
                      sum(observation[clinical_manifestation == "disease" & matching_type == 1])) %>%
      dplyr::mutate(matching_type_carriage =
                      sum(observation[clinical_manifestation == "carriage" & matching_type == 1])) %>%
      dplyr::mutate(non_matching_type_disease =
                      sum(observation[clinical_manifestation == "disease" & matching_type == 0])) %>%
      dplyr::mutate(non_matching_type_carriage =
                      sum(observation[clinical_manifestation == "carriage" & matching_type == 0])) %>%
      dplyr::ungroup() %>%
    dplyr::select(study,matching_type_disease,matching_type_carriage,non_matching_type_disease,non_matching_type_carriage) %>%
    dplyr::distinct() %>%
    dplyr::mutate(serotype = serotype) %>%
    dplyr::filter(matching_type_carriage > 0 | matching_type_disease > 0)
  return(sero_df)
}

per_serotype_estimate <- function(test_serotype, df, method_type = "FE") {
  per_serotype_out <- metafor::rma(yi = yi,
                                   vi = vi,
                                   data = df,
                                   slab = study,
                                   subset = serotype == test_serotype,
                                   method = method_type,
                                   control=list(stepadj=0.5, maxiter=10000))
  summary_df <-
    data.frame(
      "Analysis" = method_type,
      "Serotype" = test_serotype,
      "OR" = as.numeric(per_serotype_out$b),
      "OR_upper" = per_serotype_out$ci.ub,
      "OR_lower" = per_serotype_out$ci.lb,
      "tau_squared" = per_serotype_out$tau2,
      "i_squared" = per_serotype_out$I2,
      "h_squared" = per_serotype_out$H2
    )
  return(summary_df)
}

serotype_or_df <- dplyr::bind_rows(lapply(
                                   serotype_levels,
                                   get_serotype_stats,
                                   S_pneumoniae_adult_serotype)
)

serotype_effect_size_df <- metafor::escalc(measure = "OR",
                                            ai = matching_type_disease,
                                            bi = matching_type_carriage,
                                            ci = non_matching_type_disease,
                                            di = non_matching_type_carriage,
                                            slab = study,
                                            data = serotype_or_df)

per_serotype_out <- dplyr::bind_rows(
                    lapply(
                      serotype_levels,
                      per_serotype_estimate,
                      serotype_effect_size_df
                    ),
                    lapply(
                      serotype_levels,
                      per_serotype_estimate,
                      serotype_effect_size_df,
                      method_type = "REML"
                    )
)

sample_sizes <-
  best_fitting_output %>%
      dplyr::group_by(type) %>%
      dplyr::mutate(num_samples = sum(carriage+disease)) %>%
      dplyr::mutate(Sample_count = dplyr::case_when(
                        num_samples < 11 ~ "0-10",
                        num_samples < 51 ~ "11-50",
                        num_samples < 101 ~ "51-100",
                        TRUE ~ ">100"
                      )
                    ) %>%
      dplyr::ungroup() %>%
      dplyr::select(type, Sample_count) %>%
      dplyr::distinct() %>%
      dplyr::mutate(Sample_count = factor(Sample_count,
                                          levels = c("0-10",
                                                     "11-50",
                                                     "51-100",
                                                     ">100")))

# Combine the invasiveness estimates
combined_invasiveness_df <-
  dplyr::bind_rows(
    per_serotype_out %>%
      dplyr::select(Analysis, Serotype, OR, OR_upper, OR_lower) %>%
      dplyr::rename(Invasiveness = OR) %>%
      dplyr::rename(Invasiveness_upper = OR_upper) %>%
      dplyr::rename(Invasiveness_lower = OR_lower),
    best_fitting_output %>%
      dplyr::select(model_name, nu, nu_lower, nu_upper, type) %>%
      dplyr::distinct() %>%
      dplyr::rename(Analysis = model_name) %>%
      dplyr::rename(Serotype = type) %>%
      dplyr::rename(Invasiveness = nu) %>%
      dplyr::rename(Invasiveness_upper = nu_upper) %>%
      dplyr::rename(Invasiveness_lower = nu_lower)
  ) %>%
  tidyr::pivot_wider(id_cols = Serotype,
                     names_from = Analysis,
                     values_from = c(Invasiveness,Invasiveness_upper,Invasiveness_lower),
                     names_sep = "_") %>%
    dplyr::mutate("classification" = dplyr::case_when(
        Serotype %in% vaccine_types[["PCV7"]] ~ "PCV7",
        Serotype %in% vaccine_types[["PCV10"]] ~ "PCV10",
        Serotype %in% vaccine_types[["PCV13"]] ~ "PCV13",
        Serotype %in% vaccine_types[["PCV15"]] ~ "PCV15",
        Serotype %in% vaccine_types[["PCV20"]] ~ "PCV20",
        Serotype %in% vaccine_types[["PPV23"]] ~ "PPV23",
        TRUE ~ "No PCV"
      )
    ) %>%
  dplyr::left_join(sample_sizes, by = c("Serotype" = "type"))

# Plot comparison of invasiveness estimates
ggplot(combined_invasiveness_df,
       aes(x = Invasiveness_adjusted_type_specific_poisson,
           xmin = Invasiveness_lower_adjusted_type_specific_poisson,
           xmax = Invasiveness_upper_adjusted_type_specific_poisson,
           y = Invasiveness_FE,
           ymin = Invasiveness_lower_FE,
           ymax = Invasiveness_upper_FE,
           colour = classification)) +
  geom_errorbar(alpha = 0.25) +
  geom_errorbarh(alpha = 0.25) +
  geom_point() +
  scale_x_continuous(trans = "log10") +
  theme_bw() +
  scale_color_manual(values = vaccine_colours) +
  theme(legend.position = "bottom") +
  ylab("Invasiveness estimated from log odds ratio") +
  xlab("Invasivess estimate from adjusted type-specific Poisson (cases per carrier per year)") +
  facet_wrap(~Sample_count) +
  stat_cor(aes(x = Invasiveness_adjusted_type_specific_poisson,
               y = Invasiveness_FE),
           inherit.aes = FALSE,
           method = "kendall",
           cor.coef.name = "tau") +
  ggrepel::geom_label_repel(data = combined_invasiveness_df %>%
                                    dplyr::filter((Invasiveness_adjusted_type_specific_poisson < 0.0005 & Invasiveness_FE > 0) | Serotype == "46"),
                            aes(label = Serotype)
                            )

```

``` {r Plot comparison with mixed effects}

# Plot comparison of invasiveness estimates
ggplot(combined_invasiveness_df,
       aes(x = Invasiveness_adjusted_type_specific_poisson,
           xmin = Invasiveness_lower_adjusted_type_specific_poisson,
           xmax = Invasiveness_upper_adjusted_type_specific_poisson,
           y = Invasiveness_REML,
           ymin = Invasiveness_lower_REML,
           ymax = Invasiveness_upper_REML,
           colour = classification)) +
  geom_errorbar(alpha = 0.25) +
  geom_errorbarh(alpha = 0.25) +
  geom_point() +
  scale_x_continuous(trans = "log10") +
  theme_bw() +
  scale_color_manual(values = vaccine_colours) +
  theme(legend.position = "bottom") +
  ylab("Invasiveness estimated from log odds ratio") +
  xlab("Invasivess estimate from adjusted type-specific Poisson (cases per carrier per year)") +
  facet_wrap(~Sample_count) +
  stat_cor(aes(x = Invasiveness_adjusted_type_specific_poisson,
               y = Invasiveness_REML),
           inherit.aes = FALSE,
           method = "kendall",
           cor.coef.name = "tau") +
  ggrepel::geom_label_repel(data = combined_invasiveness_df %>%
                                    dplyr::filter((Invasiveness_adjusted_type_specific_poisson < 0.0005 & Invasiveness_FE > 0) | Serotype == "46"),
                            aes(label = Serotype)
                            )

```

## Comparison with carriage duration

``` {r Carriage duration comparison}

carriage_duration.df <- data.frame(
  "type" = c("19F","23F","6B","14","21","19B","18C","29","3","4","24F","NT","5","6A","6C"),
  "carriage_duration" = c(46.9,21,16.2,7.2,1.6,-0.1,-1.9,-4.3,-4.5,-7.2,-8.5,-12.3,-18.6,0,0)
)

carriage_duration.df %<>%
  dplyr::inner_join(pv_best_fitting_output,
                    by = "type") %>%
  dplyr::select(type,carriage_duration,classification,nu,nu_lower,nu_upper) %>%
  dplyr::distinct()

ggplot(carriage_duration.df,
       aes(x = carriage_duration,
           y = nu,
           ymin = nu_lower,
           ymax = nu_upper,
           colour = classification)) +
  geom_point() +
  geom_errorbar() +
  scale_colour_manual(values = vaccine_colours) +
  ggrepel::geom_label_repel(aes(label = type)) +
  scale_y_continuous(trans = "log10") +
  theme_bw()

```
