---
title: "*S. pneumoniae* strain invasiveness"
output: rmarkdown::html_vignette
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width = 8,
  comment = "#>"
)
```

```{r setup}
# Load dependencies
require(tidyverse)
require(magrittr)
require(rstan)
require(bridgesampling)
require(loo)
require(cowplot)
require(ggrepel)
require(xlsx)
require(gtools)
# Load stan package
library(progressionEstimation)
pkgbuild::compile_dll()
roxygen2::roxygenize(package.dir = "..")

# Define serotype categories and palettes
vaccine_types <- list(
  "PCV7" = c("4","6B","9V","14","18C","19F","23F"),
  "PCV10" = c("1","5","7F"),
  "PCV13" = c("6A","3","19A"),
  "PCV15" = c("22F","33F"),
  "PCV20" = c("8","10A","11A","12F","15B/C"),
  "PPV23" = c("2","9N","17F","20")
)

vaccine_colours <- c("No PCV" = "blue",
                     "PCV7" = "red",
                     "PCV10" = "orange",
                     "PCV13" = "coral2",
                     "PCV15" = "pink",
                     "PCV20" = "purple",
                     "PPV23" = "black")

epi_colors <- c(
  "carriage" = "skyblue",
  "disease" = "cornflowerblue"
)

```

## Description and filtering of the dataset

The dataset used in this analysis is contained in the `S_pneumoniae_infant_strain` object, which includes data from `r S_pneumoniae_infant_strain %>% dplyr::select(study) %>% dplyr::n_distinct()` studies on `r S_pneumoniae_infant_strain %>% dplyr::select(type) %>% dplyr::n_distinct()` serotypes. The distribution of samples between each study can be plotted, if they are ranked by the mean number of carriage and disease isolates:

``` {r Plot distribution by study, include = FALSE}

studies_by_size <-
  S_pneumoniae_infant_strain %>%
    group_by(study) %>%
    dplyr::mutate(total_count = sum(carriage)+sum(disease)) %>%
    dplyr::select(study,total_count) %>%
    dplyr::distinct() %>%
    dplyr::arrange(desc(total_count)) %>%
    dplyr::select(study) %>%
    dplyr::pull()

S_pneumoniae_infant_strain %<>%
  dplyr::mutate(study = factor(study, levels = studies_by_size))

ggplot(S_pneumoniae_infant_strain %>%
         tidyr::pivot_longer(c(carriage,disease),
                             names_to = "Sample type",
                             values_to = "count"),
       aes(x = study,
           y = count,
           colour = `Sample type`,
           fill = `Sample type`)) +
  geom_col() +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  scale_fill_manual(values = epi_colors) +
  scale_colour_manual(values = epi_colors)

```

The total number of observations for each serotype can be similarly plotted:

``` {r Plot distribution by serotype, include = FALSE}

serotype_by_count <-
  S_pneumoniae_infant_strain %>%
    group_by(type) %>%
    dplyr::mutate(total_count = sum(carriage)+sum(disease)) %>%
    dplyr::select(type,total_count) %>%
    dplyr::distinct() %>%
    dplyr::arrange(desc(total_count)) %>%
    dplyr::select(type) %>%
    dplyr::pull()

S_pneumoniae_infant_strain %<>%
  dplyr::mutate(type = factor(type, levels = serotype_by_count))

ggplot(S_pneumoniae_infant_strain %>%
         tidyr::pivot_longer(c(carriage,disease),
                             names_to = "Sample type",
                             values_to = "count"),
       aes(x = type,
           y = count,
           colour = `Sample type`,
           fill = `Sample type`)) +
  geom_col() +
  xlab("Serotype") +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  scale_fill_manual(values = epi_colors) +
  scale_colour_manual(values = epi_colors)

```

As many of these serotypes are rare, to facilitate model fitting, we exclude any serotype for which the maximum number of observations is below five:

``` {r Plot distribution by serotype in filtered dataset, include = FALSE}

threshold_count <- 5

filtered_S_pneumoniae_infant_strain <-
  S_pneumoniae_infant_strain %>%
    dplyr::filter(carriage >= threshold_count | disease >= threshold_count) 

filtered_S_pneumoniae_infant_strain %<>%
    dplyr::mutate(study = factor(study, levels = studies_by_size[studies_by_size %in% unique(filtered_S_pneumoniae_infant_strain$study)]))

filtered_S_pneumoniae_infant_strain %<>%
    dplyr::mutate(type = factor(type,
                                          levels = serotype_by_count[serotype_by_count %in% unique(filtered_S_pneumoniae_infant_strain$type)])) %>%
  dplyr::mutate(strain = factor(strain))

# Generate input for serotype models
infant_serotype_invasiveness <- 
  progressionEstimation::process_input_data(filtered_S_pneumoniae_infant_strain,
                                            condense = FALSE)

# Generate input for strain models
infant_strain_invasiveness <- 
  progressionEstimation::process_input_data(filtered_S_pneumoniae_infant_strain,
                                            type = "strain",
                                            condense = FALSE)

# Generate input for strain and serotype models
infant_serotype_and_strain_invasiveness <- 
  progressionEstimation::process_input_data(filtered_S_pneumoniae_infant_strain,
                                            use_strain = TRUE,
                                            condense = FALSE)

# Generate input for serotype models
infant_serotype_strain_combined_invasiveness <- 
  progressionEstimation::process_input_data(filtered_S_pneumoniae_infant_strain,
                                            combine_strain = TRUE,
                                            condense = FALSE)

ggplot(filtered_S_pneumoniae_infant_strain %>%
         tidyr::pivot_longer(c(carriage,disease),
                             names_to = "Sample type",
                             values_to = "count"),
       aes(x = type,
           y = count,
           colour = `Sample type`,
           fill = `Sample type`)) +
  geom_col() +
  xlab("Serotype") +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  scale_fill_manual(values = epi_colors) +
  scale_colour_manual(values = epi_colors)

```

This filtered dataset comprises `r filtered_S_pneumoniae_infant_strain %>% dplyr::select(type) %>% dplyr::n_distinct()` serotypes.

## Fitting the statistical models

An adjusted type-specific Poisson model can be applied to the data using only classification by serotype (**adjusted serotype-specific Poisson** model):

$$d_{i,j,k} \sim Poisson(\delta_{i}\nu_{j}\rho_{i,j,k}N_{i}t_{i})$$

``` {r Adjusted serotype-specific Poisson model fit, include = FALSE}

n_chains <- 2
n_iter <- 2.5e4
n_core <- 2

adjusted_serotype_specific_poisson_fit <- 
  progressionEstimation::fit_progression_rate_model(infant_serotype_invasiveness,
                                                    type_specific = TRUE,
                                                    location_adjustment = TRUE,
                                                    stat_model = "poisson",
                                                    model_description = "serotype_specific_poisson",
                                                    num_chains = n_chains,
                                                    num_iter = n_iter,
                                                    num_cores = n_core)

adjusted_serotype_specific_poisson_output_df <-
  progressionEstimation::process_progression_rate_model_output(adjusted_serotype_specific_poisson_fit,
                                                               filtered_S_pneumoniae_infant_strain,
                                                               condense = FALSE)

```

We can alternatively fit a negative binomial version of the same model:

$$d_{i,j,k} \sim NegBin(\delta_{i}\nu_{j}\rho_{i,j,k}N_{i}t_{i})$$

``` {r Adjusted serotype-specific negative binomial model fit, include = FALSE}

adjusted_serotype_specific_negbin_fit <- 
  progressionEstimation::fit_progression_rate_model(infant_serotype_invasiveness,
                                                    type_specific = TRUE,
                                                    location_adjustment = TRUE,
                                                    stat_model = "negbin",
                                                    model_description = "serotype_specific_negbin",
                                                    num_chains = n_chains,
                                                    num_iter = n_iter,
                                                    num_cores = n_core)

adjusted_serotype_specific_negbin_output_df <-
  progressionEstimation::process_progression_rate_model_output(adjusted_serotype_specific_negbin_fit,
                                                               filtered_S_pneumoniae_infant_strain,
                                                               condense = FALSE)

```


An adjusted type-specific Poisson model can alternatively be applied to the data using only classification by strain (**adjusted strain-specific Poisson** model):

$$d_{i,j,k} \sim Poisson(\delta_{i}\nu_{k}\rho_{i,j,k}N_{i}t_{i})$$

``` {r Adjusted strain-specific Poisson model fit, include = FALSE}

adjusted_strain_specific_poisson_fit <- 
  progressionEstimation::fit_progression_rate_model(infant_strain_invasiveness,
                                                    type_specific = TRUE,
                                                    location_adjustment = TRUE,
                                                    stat_model = "poisson",
                                                    model_description = "strain_specific_poisson",
                                                    num_chains = n_chains,
                                                    num_iter = n_iter,
                                                    num_cores = n_core)

adjusted_strain_specific_poisson_output_df <-
  progressionEstimation::process_progression_rate_model_output(adjusted_strain_specific_poisson_fit,
                                                               filtered_S_pneumoniae_infant_strain,
                                                               type = "strain",
                                                               condense = FALSE)

```

We can also fit a negative binomial equivalent (**adjusted strain-specific negative binomial** model):

$$d_{i,j,k} \sim NegBin(\delta_{i}\nu_{k}\rho_{i,j,k}N_{i}t_{i})$$
``` {r Adjusted strain-specific negative binomial model fit, include = FALSE}

adjusted_strain_specific_negbin_fit <- 
  progressionEstimation::fit_progression_rate_model(infant_strain_invasiveness,
                                                    type_specific = TRUE,
                                                    location_adjustment = TRUE,
                                                    stat_model = "negbin",
                                                    model_description = "strain_specific_negbin",
                                                    num_chains = n_chains,
                                                    num_iter = n_iter,
                                                    num_cores = n_core)

adjusted_strain_specific_negbin_output_df <-
  progressionEstimation::process_progression_rate_model_output(adjusted_strain_specific_negbin_fit,
                                                               filtered_S_pneumoniae_infant_strain,
                                                               type = "strain",
                                                               condense = FALSE)

```

Models can also combine serotype and strain information, with invasiveness primarily determined by serotype, and modified by strain:

$$d_{i,j,k} \sim Poisson(\delta_{i}\nu_{k}\nu_{k}\rho_{i,j,k}N_{i}t_{i})$$

``` {r Adjusted serotype-specific strain-modified Poisson model fit, include = FALSE}

adjusted_serotype_specific_strain_modified_poisson_fit <- 
  progressionEstimation::fit_progression_rate_model(infant_serotype_and_strain_invasiveness,
                                                    type_specific = TRUE,
                                                    location_adjustment = TRUE,
                                                    strain_as_secondary_type = TRUE,
                                                    stat_model = "poisson",
                                                    model_description = "serotype_specific_strain_modified_poisson",
                                                    num_chains = n_chains,
                                                    num_iter = n_iter,
                                                    num_cores = n_core)

adjusted_serotype_specific_strain_modified_poisson_output_df <-
  progressionEstimation::process_progression_rate_model_output(adjusted_serotype_specific_strain_modified_poisson_fit,
                                                               filtered_S_pneumoniae_infant_strain,
                                                               strain_as_secondary_type = TRUE,
                                                               condense = FALSE)

```

This can also be applied using a negative binomial distribution:

$$d_{i,j,k} \sim NegBin(\delta_{i}\nu_{k}\nu_{k}\rho_{i,j,k}N_{i}t_{i})$$

``` {r Adjusted serotype-specific strain-modified negative binomial model fit, include = FALSE}

adjusted_strain_specific_strain_modified_negbin_fit <- 
  progressionEstimation::fit_progression_rate_model(infant_serotype_and_strain_invasiveness,
                                                    type_specific = TRUE,
                                                    location_adjustment = TRUE,
                                                    strain_as_secondary_type = TRUE,
                                                    stat_model = "negbin",
                                                    model_description = "serotype_specific_strain_modified_negbin",
                                                    num_chains = n_chains,
                                                    num_iter = n_iter,
                                                    num_cores = n_core)

adjusted_serotype_specific_strain_modified_negbin_output_df <-
  progressionEstimation::process_progression_rate_model_output(adjusted_strain_specific_strain_modified_negbin_fit,
                                                               filtered_S_pneumoniae_infant_strain,
                                                               strain_as_secondary_type = TRUE,
                                                               condense = FALSE)

```

Models can alternative have invasiveness primarily determined by strain, and modified by serotype:

$$d_{i,j,k} \sim Poisson(\delta_{i}\nu_{k}\nu_{k}\rho_{i,j,k}N_{i}t_{i})$$

``` {r Adjusted serotype-specific strain-modified Poisson model fit, include = FALSE}

adjusted_strain_specific_serotype_modified_poisson_fit <- 
  progressionEstimation::fit_progression_rate_model(infant_serotype_and_strain_invasiveness,
                                                    type_specific = TRUE,
                                                    location_adjustment = TRUE,
                                                    strain_as_primary_type = TRUE,
                                                    stat_model = "poisson",
                                                    model_description = "strain_specific_serotype_modified_poisson",
                                                    num_chains = n_chains,
                                                    num_iter = n_iter,
                                                    num_cores = n_core)

adjusted_strain_specific_serotype_modified_poisson_output_df <-
  progressionEstimation::process_progression_rate_model_output(adjusted_strain_specific_serotype_modified_poisson_fit,
                                                               filtered_S_pneumoniae_infant_strain,
                                                               strain_as_primary_type = TRUE,
                                                               condense = FALSE)

```

This model can also be applied using a negative binomial model:

$$d_{i,j,k} \sim NegBin(\delta_{i}\nu_{k}\nu_{k}\rho_{i,j,k}N_{i}t_{i})$$

``` {r Adjusted serotype-specific strain-modified Poisson model fit, include = FALSE}

adjusted_strain_specific_serotype_modified_negbin_fit <- 
  progressionEstimation::fit_progression_rate_model(infant_serotype_and_strain_invasiveness,
                                                    type_specific = TRUE,
                                                    location_adjustment = TRUE,
                                                    strain_as_primary_type = TRUE,
                                                    stat_model = "negbin",
                                                    model_description = "strain_specific_serotype_modified_negbin",
                                                    num_chains = n_chains,
                                                    num_iter = n_iter,
                                                    num_cores = n_core)

adjusted_strain_specific_serotype_modified_negbin_output_df <-
  progressionEstimation::process_progression_rate_model_output(adjusted_strain_specific_serotype_modified_negbin_fit,
                                                               filtered_S_pneumoniae_infant_strain,
                                                               strain_as_primary_type = TRUE,
                                                               condense = FALSE)

```

The final model structure assigns a different invasiveness to each strain-serotype combination:

$$d_{i,j,k} \sim Poisson(\delta_{i}\nu_{j,k}\rho_{i,j,k}N_{i}t_{i})$$

``` {r Adjusted serotype-specific strain-modified Poisson model fit, include = FALSE}

adjusted_strain_and_serotype_specific_poisson_fit <- 
  progressionEstimation::fit_progression_rate_model(infant_serotype_strain_combined_invasiveness,
                                                    type_specific = TRUE,
                                                    location_adjustment = TRUE,
                                                    stat_model = "poisson",
                                                    model_description = "strain_and_serotype_specific_poisson",
                                                    num_chains = n_chains,
                                                    num_iter = n_iter,
                                                    num_cores = n_core)

adjusted_strain_and_serotype_specific_poisson_output_df <-
  progressionEstimation::process_progression_rate_model_output(adjusted_strain_and_serotype_specific_poisson_fit,
                                                               filtered_S_pneumoniae_infant_strain,
                                                               combined_strain_type = TRUE,
                                                               condense = FALSE)

```

A version using the negative binomial distribution can also be fitted:

$$d_{i,j,k} \sim NegBin(\delta_{i}\nu_{j,k}\rho_{i,j,k}N_{i}t_{i})$$
``` {r Adjusted serotype-specific strain-modified negative binomial model fit, include = FALSE}

adjusted_strain_and_serotype_specific_negbin_fit <- 
  progressionEstimation::fit_progression_rate_model(infant_serotype_strain_combined_invasiveness,
                                                    type_specific = TRUE,
                                                    location_adjustment = TRUE,
                                                    stat_model = "negbin",
                                                    model_description = "strain_and_serotype_specific_negbin",
                                                    num_chains = n_chains,
                                                    num_iter = n_iter,
                                                    num_cores = n_core)

adjusted_strain_and_serotype_specific_negbin_output_df <-
  progressionEstimation::process_progression_rate_model_output(adjusted_strain_and_serotype_specific_negbin_fit,
                                                               filtered_S_pneumoniae_infant_strain,
                                                               combined_strain_type = TRUE,
                                                               condense = FALSE)

```

To check the models have fitted correctly, we can plot the $\hat{R}$ values across MCMCs:

``` {r Plot r hat values}

fit_list <- list(adjusted_serotype_specific_poisson_fit,
                                  adjusted_serotype_specific_negbin_fit,
                                  adjusted_strain_specific_poisson_fit,
                                  adjusted_strain_specific_negbin_fit,
                                  adjusted_serotype_specific_strain_modified_poisson_fit,
                                  adjusted_strain_specific_strain_modified_negbin_fit,
                                  adjusted_strain_specific_serotype_modified_poisson_fit,
                                  adjusted_strain_specific_serotype_modified_negbin_fit,
                                  adjusted_strain_and_serotype_specific_poisson_fit,
                                  adjusted_strain_and_serotype_specific_negbin_fit)

rhat_plot <- function(fit) {
  rhat_plot <- plot(fit, plotfun = "rhat", binwidth = 0.00005)
  rhat_plot <- rhat_plot + theme(axis.text.x = element_text(angle = 90))
}

rhat_plots <- lapply(fit_list, rhat_plot)

(strain_and_serotype_rhat_plots <-
  cowplot::plot_grid(plotlist = rhat_plots, ncol = 5, nrow = 2, labels = "AUTO")
)

ggsave(strain_and_serotype_rhat_plots,
       file = "strain_and_serotype_rhat_plots.png",
       width = 11,
       height = 7
)

```

Convergence can also be checked using trace plots of the log likelihood values:

``` {r Plot log likelihood traces}

plot_lp <- function(fit) {
  pl <- rstan::traceplot(fit, pars = "lp__")
  no_leg_pl <- pl + theme(legend.position = "none",
                          axis.text.x = element_text(angle = 90))
  return(no_leg_pl)
}

trace_plots <- lapply(fit_list, plot_lp)

(strain_and_serotype_trace_plots <-
  cowplot::plot_grid(plotlist = trace_plots, ncol = 5, nrow = 2, labels = "AUTO")
)

ggsave(strain_and_serotype_trace_plots,
       file = "strain_and_serotype_trace_plots.png",
       width = 11,
       height = 7
)

```

## Compare model observations and predictions

For each model, the output can be processed, and the observations plotted against each model's predictions.

``` {r Process model outputs and plot predictions, fig.height = 18}

model_outputs <- list(
  adjusted_serotype_specific_poisson_output_df,
  adjusted_serotype_specific_negbin_output_df,
  adjusted_strain_specific_poisson_output_df,
  adjusted_strain_specific_negbin_output_df,
  adjusted_serotype_specific_strain_modified_poisson_output_df,
  adjusted_serotype_specific_strain_modified_negbin_output_df,
  adjusted_strain_specific_serotype_modified_poisson_output_df,
  adjusted_strain_specific_serotype_modified_negbin_output_df,
  adjusted_strain_and_serotype_specific_poisson_output_df,
  adjusted_strain_and_serotype_specific_negbin_output_df
)

label_list <- list(
  "type",
  "type",
  "type",
  "type",
  "type",
  "type",
  "type",
  "type",
  "combined",
  "combined"
)

prediction_plots <- list()
for (i in 1:length(model_outputs)) {
  prediction_plots[[i]] <- progressionEstimation::plot_case_carrier_predictions(
                                                       model_outputs[[i]],
                                                       label_col = label_list[[i]],
                                                       legend = FALSE)
}

no_leg_plot <- cowplot::plot_grid(plotlist = prediction_plots, ncol = 1, nrow = 10, labels = "AUTO")

prediction_plot_legend <- progressionEstimation::plot_case_carrier_predictions(model_outputs[[1]], just_legend = TRUE)

cowplot::plot_grid(no_leg_plot, prediction_plot_legend, ncol = 1, rel_heights = c(0.9,0.1))

```

The deviation of the predicted and observed values can be simply evaluated using the root mean square error:

``` {r Calculate RMSE}

obs_pred_df <-
  dplyr::bind_rows(model_outputs) %>%
  dplyr::select(model_name,carriage_rmse,disease_rmse,phi,phi_lower,phi_upper) %>%
  tidyr::pivot_longer(c(carriage_rmse,disease_rmse), names_to = "Observation", values_to = "RMSE") %>%
  dplyr::mutate(model_name = factor(model_name,
                                    levels = c(
                                      "serotype_specific_poisson",
                                      "serotype_specific_negbin",
                                      "strain_specific_poisson",
                                      "strain_specific_negbin",
                                      "serotype_specific_strain_modified_poisson",
                                      "serotype_specific_strain_modified_negbin",
                                      "strain_specific_serotype_modified_poisson",
                                      "strain_specific_serotype_modified_negbin",
                                      "strain_and_serotype_specific_poisson",
                                      "strain_and_serotype_specific_negbin"
                                      )
                                    )
                ) %>%
  dplyr::mutate(Observation = dplyr::case_when(
                    Observation == "carriage_rmse" ~ "Carriage",
                    Observation == "disease_rmse" ~ "Disease"
                  )
                )

(strain_and_serotype_rmse_plot <-
  ggplot(obs_pred_df, aes(x = model_name, y = RMSE)) +
    geom_point(position = position_jitter(width = 0.25),
               alpha = 0.25,
               cex = 0.75,
               colour = "blue",
               pch = 4) +
    geom_violin(fill = NA) +
    scale_y_continuous(trans="log10") +
    xlab("Model name") +
    theme_bw() +
    facet_grid(Observation ~ .) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
)

ggsave(strain_and_serotype_rmse_plot,
       file = "strain_and_serotype_rmse_plot.png",
       width = 7,
       height = 7
)

```

The fits of the negative binomial can be evaluated through plotting the estimates of the precision parameter, $\phi$. As $\phi$ increases, the variance of the negative binomial distribution decreases, and the negative binomal model becomes more similar to the Poisson model.

``` {r Plot precision parameters for negative binomial models, fig.width = 6}

(strain_and_serotype_precision_parameters <-
  ggplot(obs_pred_df %>%
          dplyr::filter(!is.na(phi)) %>%
          dplyr::group_by(model_name) %>%
          dplyr::slice(1) %>%
          dplyr::ungroup(),
        aes(x = model_name,
            y = phi,
            ymin = phi_lower,
            ymax = phi_upper)) +
    geom_point() +
    geom_errorbar(width = 0.25) +
    theme_bw() +
    ylab(expression(phi)) +
    xlab("Model name") +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
)

ggsave(strain_and_serotype_precision_parameters,
       file = "strain_and_serotype_precision_plot.png",
       width = 7,
       height = 7
)

```

## Run model comparison with leave-one-out cross-validation

The model likelihoods can be compared using leave-one-out cross-validation:

``` {r Comparison with LOO-CV using full log likelihood, include = FALSE}

loo_comparison <- progressionEstimation::compare_model_fits_with_loo(fit_list,
                                                                     use_moments = FALSE,
                                                                     num_threads = n_core)

loo_comparison %>%
  kableExtra::kable()  %>%
  kableExtra::kable_styling(latex_options = "scale_down")

```

The Pareto distribution k values are too high for robust estimates of the ELPD. This reflects many of the points being 

``` {r Comparison with LOO-CV using disease distribution log likelihood, include = FALSE}

disease_loo_comparison <- progressionEstimation::compare_model_fits_with_loo(fit_list,
                                                                     log_lik_param = "disease_log_lik",
                                                                     use_moments = FALSE,
                                                                     num_threads = n_core)

disease_loo_comparison %>%
  kableExtra::kable()  %>%
  kableExtra::kable_styling(latex_options = "scale_down")

```

## Run model comparison with Bayes factors

As the models can instead be compared with Bayes factors:

``` {r Comparison with BF, include = FALSE}

bf_comparison <- progressionEstimation::compare_model_fits_with_bf(fit_list,
                                                                   num_iter = n_iter,
                                                                   num_threads = n_core)
bf_comparison %>%
  kableExtra::kable()  %>%
  kableExtra::kable_styling(latex_options = "scale_down")

```

## Plot serotype invasiveness

The best-fitting model can then be applied to the overall dataset.

``` {r Plot invasiveness estimates from best-fitting model}

full_infant_serotype_and_strain_invasiveness <- 
  progressionEstimation::process_input_data(S_pneumoniae_infant_strain,
                                            combine_strain = TRUE,
                                            condense = FALSE)

full_adjusted_strain_and_serotype_specific_poisson_fit <- 
  progressionEstimation::fit_progression_rate_model(full_infant_serotype_and_strain_invasiveness,
                                                    type_specific = TRUE,
                                                    location_adjustment = TRUE,
                                                    stat_model = "poisson",
                                                    model_description = "strain_and_serotype_specific_poisson",
                                                    num_chains = n_chains,
                                                    num_iter = n_iter,
                                                    num_cores = n_core)

rhat_plot <- plot(full_adjusted_strain_and_serotype_specific_poisson_fit, plotfun = "rhat", binwidth = 0.00005)

trace_plot <- rstan::traceplot(full_adjusted_strain_and_serotype_specific_poisson_fit, pars = "lp__")

(full_strain_and_serotype_validation_plot <-
  cowplot::plot_grid(plotlist = list(rhat_plot,trace_plot),
                     nrow = 1,
                     ncol = 2,
                     labels = "AUTO")
)

ggsave(full_strain_and_serotype_validation_plot,
       file = "full_strain_and_serotype_validation_plot.png",
       width = 7,
       height = 5
)

```

``` {r Plot invasiveness across all serotypes}

# Annotate serotypes by status
best_fitting_output <- 
  progressionEstimation::process_progression_rate_model_output(full_adjusted_strain_and_serotype_specific_poisson_fit,
                                                               S_pneumoniae_infant_strain,
                                                               combined_strain_type = TRUE) %>%
    dplyr::mutate("type" = S_pneumoniae_infant_strain$type) %>%
    dplyr::mutate("classification" = dplyr::case_when(
        type %in% vaccine_types[["PCV7"]] ~ "PCV7",
        type %in% vaccine_types[["PCV10"]] ~ "PCV10",
        type %in% vaccine_types[["PCV13"]] ~ "PCV13",
        type %in% vaccine_types[["PCV15"]] ~ "PCV15",
        type %in% vaccine_types[["PCV20"]] ~ "PCV20",
        type %in% vaccine_types[["PPV23"]] ~ "PPV23",
        TRUE ~ "No PCV"
      )
    ) %>%
    dplyr::mutate(classification = factor(classification, levels = c("PCV7",
                                                                     "PCV10",
                                                                     "PCV13",
                                                                     "PCV15",
                                                                     "PCV20",
                                                                     "PPV23",
                                                                     "No PCV")))

# Order serotypes
strains_by_invasiveness <-
  best_fitting_output %>%
    dplyr::select(combined,nu) %>%
    dplyr::distinct() %>%
    dplyr::arrange(desc(nu)) %>%
    dplyr::select(combined) %>%
    dplyr::pull()

best_fitting_output %<>%
  dplyr::mutate(combined = factor(combined, levels = strains_by_invasiveness))

# Plot
(full_invasiveness_plot <-
  progressionEstimation::plot_progression_rates(best_fitting_output,
                                                type = "combined",
                                                unit_time = "year",
                                                type_name = "Serotype",
                                                colour_col = "classification",
                                                colour_palette = vaccine_colours,
                                                use_sample_size = TRUE) +
    theme(axis.text.x = element_text(size = 6))
)

ggsave(full_invasiveness_plot,
       file = "full_strain_and_serotype_invasiveness_plot.png",
       width = 20,
       height = 10
)

```



``` {r Plot only larger sample size}

# Annotate sample size
best_fitting_output %<>%
  dplyr::group_by(combined) %>%
  dplyr::mutate(num_observations = sum(carriage+disease)) %>%
  dplyr::ungroup() %>%
  dplyr::filter(num_observations >= 10) %>%
  dplyr::group_by(strain) %>%
  dplyr::mutate(num_serotypes = dplyr::n_distinct(type)) %>%
  dplyr::ungroup() %>%
  dplyr::group_by(type) %>%
  dplyr::mutate(num_strains = dplyr::n_distinct(strain)) %>%
  dplyr::ungroup()

# Plot
(invasiveness_plot <-
  progressionEstimation::plot_progression_rates(best_fitting_output,
                                                type = "combined",
                                                unit_time = "year",
                                                type_name = "Serotype",
                                                colour_col = "classification",
                                                colour_palette = vaccine_colours,
                                                use_sample_size = TRUE))

```

``` {r Plot variation within serotypes, fig.height = 12}

mixedrank = function(x) order(gtools::mixedorder(as.character(x)))

# Reorder by serotype
strain_and_serotype_levels <-
  best_fitting_output %>%
    dplyr::select(combined,classification) %>%
    dplyr::distinct() %>%
    dplyr::arrange(classification,combined) %>%
    dplyr::select(combined) %>%
    dplyr::pull()

serotype_levels <-
  best_fitting_output %>%
    dplyr::select(type,classification) %>%
    dplyr::distinct() %>%
    dplyr::group_by(classification) %>%
    dplyr::arrange(mixedrank(type),
                   .by_group = TRUE) %>%
    dplyr::ungroup() %>%
    dplyr::select(type) %>%
    dplyr::pull()

best_fitting_output %<>%
  dplyr::mutate(combined = factor(combined, levels = strain_and_serotype_levels)) %>%
  dplyr::mutate(type = factor(type, levels = serotype_levels))

(serotype_invasiveness_plot <-
  progressionEstimation::plot_progression_rates(best_fitting_output %>%
                                                  dplyr::filter(num_strains > 1),
                                                type = "combined",
                                                unit_time = "year",
                                                type_name = "Serotype",
                                                colour_col = "classification",
                                                colour_palette = vaccine_colours,
                                                use_sample_size = TRUE) +
                                                facet_wrap(.~type,
                                                           scales = 'free_x'))

ggsave(serotype_invasiveness_plot,
       file = "within_serotype_invasiveness_variation_plot.png",
       width = 8,
       height = 12
)

```


``` {r Plot variation within strains, fig.height = 12}

strain_levels <-
  best_fitting_output %>%
    dplyr::select(strain) %>%
    dplyr::distinct() %>%
    dplyr::arrange(mixedrank(strain)) %>%
    dplyr::pull()

best_fitting_output %<>%
  dplyr::mutate(strain = factor(strain, levels = strain_levels))

(strain_invasiveness_plot <-
  progressionEstimation::plot_progression_rates(best_fitting_output %>%
                                                  dplyr::filter(num_serotypes > 1),
                                                type = "combined",
                                                unit_time = "year",
                                                type_name = "Serotype",
                                                colour_col = "classification",
                                                colour_palette = vaccine_colours,
                                                use_sample_size = TRUE) +
                                                facet_wrap(.~strain,
                                                           scales = 'free_x'))

ggsave(strain_invasiveness_plot,
       file = "within_strain_invasiveness_variation_plot.png",
       width = 8,
       height = 8
)

```

## Plot study-specific scale factors

``` {r Plot study-specific scale factor estimates from best-fitting model}

(scale_factor_plot <-
  progressionEstimation::plot_study_scale_factors(best_fitting_output))

ggsave(scale_factor_plot,
       file = "full_strain_and_serotype_scale_factor_plot.png",
       width = 6,
       height = 6
)

```

``` {r Sensitivity to population sizes}

sensitivity_S_pneumoniae_infant_strain <-
  S_pneumoniae_infant_strain %>%
    dplyr::filter(carriage >= threshold_count | disease >= threshold_count) %>%
    dplyr::mutate(carriage_samples = dplyr::case_when(
                      study %in% c("Finland.Pre.PCV","Oxford.pre.PCV","Stockholm.pre.PCV") ~ as.integer(carriage_samples*100),
                      TRUE ~ carriage_samples
                    )
                  )

sensitivity_S_pneumoniae_infant_strain %<>%
    dplyr::mutate(study = factor(study, levels = studies_by_size[studies_by_size %in% unique(sensitivity_S_pneumoniae_infant_strain$study)]))

sensitivity_S_pneumoniae_infant_strain %<>%
    dplyr::mutate(type = factor(type,
                                          levels = serotype_by_count[serotype_by_count %in% unique(sensitivity_S_pneumoniae_infant_strain$type)])) %>%
  dplyr::mutate(strain = factor(strain))

# Generate input for serotype models
sensitivity_infant_serotype_invasiveness <- 
  progressionEstimation::process_input_data(sensitivity_S_pneumoniae_infant_strain,
                                            condense = FALSE)

# Generate input for strain models
sensitivity_infant_strain_invasiveness <- 
  progressionEstimation::process_input_data(sensitivity_S_pneumoniae_infant_strain,
                                            type = "strain",
                                            condense = FALSE)

# Generate input for strain and serotype models
sensitivity_infant_serotype_and_strain_invasiveness <- 
  progressionEstimation::process_input_data(sensitivity_S_pneumoniae_infant_strain,
                                            use_strain = TRUE,
                                            condense = FALSE)

# Generate input for serotype models
sensitivity_infant_serotype_strain_combined_invasiveness <- 
  progressionEstimation::process_input_data(sensitivity_S_pneumoniae_infant_strain,
                                            combine_strain = TRUE,
                                            condense = FALSE)


```

Fit the first model:

``` {r First sensitivity test}
sensitivity_adjusted_serotype_specific_poisson_fit <- 
  progressionEstimation::fit_progression_rate_model(sensitivity_infant_serotype_invasiveness,
                                                    type_specific = TRUE,
                                                    location_adjustment = TRUE,
                                                    stat_model = "poisson",
                                                    model_description = "serotype_specific_poisson",
                                                    num_chains = n_chains,
                                                    num_iter = n_iter,
                                                    num_cores = n_core)

sensitivity_adjusted_serotype_specific_poisson_output_df <-
  progressionEstimation::process_progression_rate_model_output(sensitivity_adjusted_serotype_specific_poisson_fit,
                                                               sensitivity_S_pneumoniae_infant_strain,
                                                               condense = FALSE)
```

``` {r Adjusted serotype-specific negative binomial model fit sensitivity test, include = FALSE}

sensitivity_adjusted_serotype_specific_negbin_fit <- 
  progressionEstimation::fit_progression_rate_model(sensitivity_infant_serotype_invasiveness,
                                                    type_specific = TRUE,
                                                    location_adjustment = TRUE,
                                                    stat_model = "negbin",
                                                    model_description = "serotype_specific_negbin",
                                                    num_chains = n_chains,
                                                    num_iter = n_iter,
                                                    num_cores = n_core)

sensitivity_adjusted_serotype_specific_negbin_output_df <-
  progressionEstimation::process_progression_rate_model_output(sensitivity_adjusted_serotype_specific_negbin_fit,
                                                               sensitivity_S_pneumoniae_infant_strain,
                                                               condense = FALSE)

```

``` {r Adjusted strain-specific Poisson model fit sensitivity test, include = FALSE}

sensitivity_adjusted_strain_specific_poisson_fit <- 
  progressionEstimation::fit_progression_rate_model(sensitivity_infant_strain_invasiveness,
                                                    type_specific = TRUE,
                                                    location_adjustment = TRUE,
                                                    stat_model = "poisson",
                                                    model_description = "strain_specific_poisson",
                                                    num_chains = n_chains,
                                                    num_iter = n_iter,
                                                    num_cores = n_core)

sensitivity_adjusted_strain_specific_poisson_output_df <-
  progressionEstimation::process_progression_rate_model_output(sensitivity_adjusted_strain_specific_poisson_fit,
                                                               sensitivity_S_pneumoniae_infant_strain,
                                                               type = "strain",
                                                               condense = FALSE)

```


``` {r Adjusted strain-specific negative binomial model fit sensitivity test, include = FALSE}

sensitivity_adjusted_strain_specific_negbin_fit <- 
  progressionEstimation::fit_progression_rate_model(sensitivity_infant_strain_invasiveness,
                                                    type_specific = TRUE,
                                                    location_adjustment = TRUE,
                                                    stat_model = "negbin",
                                                    model_description = "strain_specific_negbin",
                                                    num_chains = n_chains,
                                                    num_iter = n_iter,
                                                    num_cores = n_core)

sensitivity_adjusted_strain_specific_negbin_output_df <-
  progressionEstimation::process_progression_rate_model_output(sensitivity_adjusted_strain_specific_negbin_fit,
                                                               sensitivity_S_pneumoniae_infant_strain,
                                                               type = "strain",
                                                               condense = FALSE)

```






``` {r Adjusted serotype-specific strain-modified Poisson model fit sensitivity test, include = FALSE}

sensitivity_adjusted_serotype_specific_strain_modified_poisson_fit <- 
  progressionEstimation::fit_progression_rate_model(sensitivity_infant_serotype_and_strain_invasiveness,
                                                    type_specific = TRUE,
                                                    location_adjustment = TRUE,
                                                    strain_as_secondary_type = TRUE,
                                                    stat_model = "poisson",
                                                    model_description = "serotype_specific_strain_modified_poisson",
                                                    num_chains = n_chains,
                                                    num_iter = n_iter,
                                                    num_cores = n_core)

sensitivity_adjusted_serotype_specific_strain_modified_poisson_output_df <-
  progressionEstimation::process_progression_rate_model_output(sensitivity_adjusted_serotype_specific_strain_modified_poisson_fit,
                                                               sensitivity_S_pneumoniae_infant_strain,
                                                               strain_as_secondary_type = TRUE,
                                                               condense = FALSE)

```



``` {r Adjusted serotype-specific strain-modified negative binomial model fit sensitivity test, include = FALSE}

sensitivity_adjusted_strain_specific_strain_modified_negbin_fit <- 
  progressionEstimation::fit_progression_rate_model(sensitivity_infant_serotype_and_strain_invasiveness,
                                                    type_specific = TRUE,
                                                    location_adjustment = TRUE,
                                                    strain_as_secondary_type = TRUE,
                                                    stat_model = "negbin",
                                                    model_description = "serotype_specific_strain_modified_negbin",
                                                    num_chains = n_chains,
                                                    num_iter = n_iter,
                                                    num_cores = n_core)

sensitivity_adjusted_serotype_specific_strain_modified_negbin_output_df <-
  progressionEstimation::process_progression_rate_model_output(sensitivity_adjusted_strain_specific_strain_modified_negbin_fit,
                                                               sensitivity_S_pneumoniae_infant_strain,
                                                               strain_as_secondary_type = TRUE,
                                                               condense = FALSE)

```



``` {r Adjusted serotype-specific strain-modified Poisson model fit sensitivity test, include = FALSE}

sensitivity_adjusted_strain_specific_serotype_modified_poisson_fit <- 
  progressionEstimation::fit_progression_rate_model(sensitivity_infant_serotype_and_strain_invasiveness,
                                                    type_specific = TRUE,
                                                    location_adjustment = TRUE,
                                                    strain_as_primary_type = TRUE,
                                                    stat_model = "poisson",
                                                    model_description = "strain_specific_serotype_modified_poisson",
                                                    num_chains = n_chains,
                                                    num_iter = n_iter,
                                                    num_cores = n_core)

sensitivity_adjusted_strain_specific_serotype_modified_poisson_output_df <-
  progressionEstimation::process_progression_rate_model_output(sensitivity_adjusted_strain_specific_serotype_modified_poisson_fit,
                                                               sensitivity_S_pneumoniae_infant_strain,
                                                               strain_as_primary_type = TRUE,
                                                               condense = FALSE)

```



``` {r Adjusted serotype-specific strain-modified Poisson model fit sensitivity test, include = FALSE}

sensitivity_adjusted_strain_specific_serotype_modified_negbin_fit <- 
  progressionEstimation::fit_progression_rate_model(sensitivity_infant_serotype_and_strain_invasiveness,
                                                    type_specific = TRUE,
                                                    location_adjustment = TRUE,
                                                    strain_as_primary_type = TRUE,
                                                    stat_model = "negbin",
                                                    model_description = "strain_specific_serotype_modified_negbin",
                                                    num_chains = n_chains,
                                                    num_iter = n_iter,
                                                    num_cores = n_core)

sensitivity_adjusted_strain_specific_serotype_modified_negbin_output_df <-
  progressionEstimation::process_progression_rate_model_output(sensitivity_adjusted_strain_specific_serotype_modified_negbin_fit,
                                                               sensitivity_S_pneumoniae_infant_strain,
                                                               strain_as_primary_type = TRUE,
                                                               condense = FALSE)

```


``` {r Adjusted serotype-specific strain-modified Poisson model fit sensitivity test, include = FALSE}

sensitivity_adjusted_strain_and_serotype_specific_poisson_fit <- 
  progressionEstimation::fit_progression_rate_model(sensitivity_infant_serotype_strain_combined_invasiveness,
                                                    type_specific = TRUE,
                                                    location_adjustment = TRUE,
                                                    stat_model = "poisson",
                                                    model_description = "strain_and_serotype_specific_poisson",
                                                    num_chains = n_chains,
                                                    num_iter = n_iter,
                                                    num_cores = n_core)

sensitivity_adjusted_strain_and_serotype_specific_poisson_output_df <-
  progressionEstimation::process_progression_rate_model_output(sensitivity_adjusted_strain_and_serotype_specific_poisson_fit,
                                                               sensitivity_S_pneumoniae_infant_strain,
                                                               combined_strain_type = TRUE,
                                                               condense = FALSE)

```


``` {r Adjusted serotype-specific strain-modified negative binomial model fit sensitivity test, include = FALSE}

sensitivity_adjusted_strain_and_serotype_specific_negbin_fit <- 
  progressionEstimation::fit_progression_rate_model(sensitivity_infant_serotype_strain_combined_invasiveness,
                                                    type_specific = TRUE,
                                                    location_adjustment = TRUE,
                                                    stat_model = "negbin",
                                                    model_description = "strain_and_serotype_specific_negbin",
                                                    num_chains = n_chains,
                                                    num_iter = n_iter,
                                                    num_cores = n_core)

sensitivity_adjusted_strain_and_serotype_specific_negbin_output_df <-
  progressionEstimation::process_progression_rate_model_output(sensitivity_adjusted_strain_and_serotype_specific_negbin_fit,
                                                               sensitivity_S_pneumoniae_infant_strain,
                                                               combined_strain_type = TRUE,
                                                               condense = FALSE)

```

``` {r Compare sensitivity fits}

sensitivity_fit_list <- list(
  sensitivity_adjusted_serotype_specific_poisson_fit,
  sensitivity_adjusted_serotype_specific_negbin_fit,
  sensitivity_adjusted_strain_specific_poisson_fit,
  sensitivity_adjusted_strain_specific_negbin_fit,
  sensitivity_adjusted_serotype_specific_strain_modified_poisson_fit,
  sensitivity_adjusted_strain_specific_strain_modified_negbin_fit,
  sensitivity_adjusted_strain_specific_serotype_modified_poisson_fit,
  sensitivity_adjusted_strain_specific_serotype_modified_negbin_fit,
  sensitivity_adjusted_strain_and_serotype_specific_poisson_fit,
  sensitivity_adjusted_strain_and_serotype_specific_negbin_fit
)

sensitivity_bf_comparison <- 
  progressionEstimation::compare_model_fits_with_bf(
                                                    sensitivity_fit_list,
                                                    num_iter = n_iter,
                                                    num_threads = n_core)

sensitivity_bf_comparison %>%
  kableExtra::kable()  %>%
  kableExtra::kable_styling(latex_options = "scale_down")

```
